\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-04-26 forwards-and-backwards.tex
notes after meeting

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{C}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels

@@

@@

\begin{proposition}
    \label{prop:nonlocal_to_local}
    Suppose that $\eta_\epsilon(t,dx)$ is a solution to equation \eqref{eqn:nonlocal_pde}
    with $r(x, m) = 1$ for all $x$ and $m$
    and some $\epsilon_r < \epsilon_\gamma = \epsilon_F = \epsilon$
    \comment{and assumptions on $\gamma$? and $F$?}.
    Then $\eta_\epsilon(t, dx)$ converges \comment{in what sense?} as $\epsilon \to 0$
    to $u(t, x) dx$, which solves
    \begin{align} \label{eqn:PDE}
        \partial_t u(t, x)
        =
        \DG \left( \gamma(\cdot, u(t, \cdot) u(t, \cdot)  \right)(x)
        + F(x, u(t, x)) u(t, x) .
    \end{align}
\end{proposition}

@@

\begin{theorem} \label{thm:lineages}
    In the context of Definition~\ref{def:lineage_generator},
    \begin{align}
        \label{eqn:lineage_generator}
        \Lgen_s f
        &=
        \frac{r}{\varphi_s}
        \left[
            \DG^*(\gamma \varphi_s f) 
            - f \DG^* (\gamma \varphi_s)
        \right] \\
        &= \label{eqn:lineage_generator2}
        r\gamma
        \left[
            \sum_{ij} \covq_{ij} \partial_{ij} f
            + 2 \sum_j \vec{m}_j \partial_j f
        \right] ,
    \end{align}
    where $\vec{m}$ is the vector
    $$
    \vec{m}_j
    =
    2 \sum_i C_{ij} \partial_i \log(\gamma \varphi_s)
    + 2 \sum_i \partial_i C_{ij}
    - \meanq_j .
    $$
\end{theorem}

@@

Although our proof of Theorem~\ref{thm:lineages}
uses the explicit representation in terms of the lookdown process,
informal calculations agree with the main result.
Suppose that we are tracking a lineage that is at location $y$ at time $t$.
Looking back through time, the lineage will move at the time that individual was born,
and will jump to the location of the parent of that individual.
Now, the rate at which new individuals are born to parents at $x$ and establish at $y$
is
$$
    \gamma(x, \eta_t) q(x, dy) r(y, \eta_t) .
$$
This implies that the probability a randomly chosen individual near $y$
is a new offspring from a parent at $x$ in $[t, t+dt)$ is, informally,
\begin{equation} \label{eqn:informal_rates}
    \gamma(x, \eta_t) q_\theta(x, dy) r(y, \eta_t) \eta_t(dx)
    \eta_t(dy)
} dt .
Leaving aside questions of whether a lineage can be treated as a randomly chosen individual,
we might then proceed by defining a continuous-time jump process
whose transition rates are given by \eqref{eqn:informal_rates} given $(\eta_t)_{t=0}^T$.
Happily, we are interested in the case when $\eta$ has a density,
say $\eta_t(dx) = \varphi_t(x) dx$,
so interpreting these rates is not too difficult.
Now recall that $q_\theta(x, dy)$ is a Gaussian kernel
with mean $x + \meanq/\theta$ and covariance $\covq/\theta$.
Write $\widetilde q_\theta(y, dx)$ for the ``reverse-time'' distribution,
i.e., Gaussian with mean $x - \meanq/\theta$ and the same covariance.
Then, rearranging infinitesimals,
\eqref{eqn:informal_rates} implies a rate of
\begin{equation} \label{eqn:informal_rates2}
    \gamma(x, \eta_t) r(y, \eta_t) \varphi_t(x) 
    \varphi_t(y)
} \widetilde q_\theta(y, dx) .
Let $(L^\theta_s)_{s=0}^T$ be the location of a lineage that moves according to
jump rates~\eqref{eqn:informal_rates2},
back through time (i.e., with the substitution that $s = T-t$).
Then
\begin{align} \label{eqn:lineage_generator_Edt}
    \begin{split}
    &\IE[f(L^\theta_{s+ds}) \;|\; L_s = y] - f(y)
    \\&\qquad 
    ds \int \left(f(x) - f(y)\right)
        \varphi_{T-s}(x) \gamma(x, \eta_{T-s}) r(y, \eta_{T-s})
        \varphi_{T-s}(y)
    \widetilde q_\theta(y, dx) .
    \end{split}
Referring back to Definition~\ref{def:dispersal_generator},
recall that as $\theta \to \infty$,
    \theta \int (f(x) - f(y)) \widetilde q_\theta(y, dx)
    &\to
    - \sum_i \meanq_i \partial_{x_i} f(y)
    + \sum_{ij} \covq_{ij} \partial_{x_i x_j} f(y) \\
    = \DD - \meanq \cdot \grad f ,
where $\DD = \sum_{ij} \covq_{ij} \partial_{ij}$.
(Note that this differs from $\DG^*$ if $\meanq$ can vary across space.)
Consequently,
    \theta \int (f(x) - f(y)) g(x) \widetilde q_\theta(y, dx) \\
    &\qquad =
    \theta \int \left\{
        (f(x) g(x) - f(y) g(y)) - f(y) (g(x) - g(y))
    \right\} \widetilde q_\theta(y, dx) \\
    &\qquad \to
        \DD(fg)(x) - f(x) \DD g(x) - \meanq \cdot \left( \grad(fg)(x) - f(x) \grad g(x) \right) \\
    &\qquad =
        g(x) \DD f(x) + 2 (\covq \grad g(x)) \cdot \grad f(x) - g(x) \meanq \cdot \grad f(x) .
Applying this to~\eqref{eqn:lineage_generator_Edt} with $g = \varphi_{T-s} \gamma$,
this suggests that the generator of the limiting process is
\begin{align} \label{eqn:heuristic_lineage_generator}
    \Lgen_s f
    &=
    \frac{r}{\varphi_{T-s}}
        \varphi_{T-s} \gamma (\DD f - \meanq \cdot \grad f)
        + 2 \left(\covq \grad (\varphi_{T-s} \gamma) \right) \cdot \grad f 
    \right\} \\
    &=
    r \gamma
        \DD f - \meanq \cdot \grad f
        + 2 \left(\covq \grad \log(\varphi_{T-s} \gamma) \right) \cdot \grad f 
    \right\} .
In other words, the lineage moves at speed equal to the local production of successful offspring,
$r \gamma$,
by a time-reversed version of the dispersal distribution,
but with a drift in the direction of the gradient of the logarithm
of the production of juveniles, $2 \grad \log(\phi_{T-s} \gamma)$
(modified by the covariance of dispersal).
This \comment{presumably?} agrees with \comment{the theorem}.

@@

@@

\comment{TODO: consider whether this is where we stick in
    these explicit definitions of $\gamma$ and set $\lambda=N$.}
To scale our lookdown model to be consistent with
the spatial model described in \S \ref{sec: Model and main results},
we will let $\lambda$ takes on the role of $N$
in \S \ref{sec: Model and main results},
i.e. each individual has mass $1/\lambda$
in the scaling limit.
Time is once again scaled by $\theta$,
and the birth, death and establishment rates 
depend on local densities given by
mollification over Gaussian kernels with radius
$\epsilon_{\gamma}$, $\epsilon_{F}$ and $\epsilon_{r}$.

@@

\comment{This is a rewrite using Ulam-Harris labeling
of \ref{sec: Evolution of an Individual Particle}.}

@@

Concretely, then: for each label $a$ in
$\labelspace = \bigcup_{k \ge 1} \IN_{\ge 1}^k$
let $\Pi_a$ be an independent Poisson process on $[0, \infty)^2 \times \IR_d \times \{0,1\}$.
The mean measure is a product of Lesbegue measure on the first two coordinates,
$q(0, \cdot)$ on $\IR^d$, and $(\delta_0 + \delta_1)/2$ on $\{0, 1\}$.
It will also be convenient
to suppose that for each label $a$ we have an enumeration of the points in $\Pi_a$,
so we may refer to ``the $j^\text{th}$ point in $\Pi_a$'',
although the precise order of this enumeration is irrelevant.
If $(t, v, y, \kappa)$ is the $j^\text{th}$ point in $\Pi_a$,
then $t$ will determine possible birth times,
$v$ will determine levels of the offspring,
$y$ will give the spatial displacement of the offspring relative to the parent,
$\kappa$ will be used to determine whether parent or offspring are assigned the new level,
and the new label produced, if any will be $aj$,
i.e., the label $a$ with $j$ appended (so, if $a \in \IN_{\ge 0}^k$ then $aj \in \IN_{\ge 0}^{k+1}$).
Each label $a$ has a birth time $\tau_a$,
when it is first assigned,
and a death time $\sigma_a$, when its level first hits $\lambda$.
For any $\tau_a \le t \le \sigma_a$ we denote by $X_a(t)$ and $U_a(t)$ the spatial location and level
of the individual carrying label $a$ at time $t$.
Now, since we have defined labels so that the level does not jump,
$U_a$ satistifies~\eqref{eqn:dot_u} for $\tau_a \le t \le \sigma_a$, i.e.,
\begin{equation} \label{eqn:U_line_of_descent}
    \begin{split}
& U_a(t)
    =
    U_a(\tau_a) \\
&\qquad {}   
    + \int_{\tau_a}^{t}
    \left(
        \frac{\theta}{\lambda} \gamma(X_a(s),\eta(s))
        \int_{\IR^d} r(z,\eta(s)) q(X_a(s),dz) U_a(s)^2
        -
        b(X_a(s),\eta(s)) U_a(s)
    \right)
    ds ,
\end{split}
\end{equation}
and, of course, $\sigma_a = \inf\{t \ge \tau_a : U_a(t) > \lambda\}$.

Potential reproduction events occur at times $\tau_a + s$
for each point $(s, v, y, \kappa) \in \Pi_a$.
We say `potential' since if the resulting level is greater than $\lambda$,
the event does not happen.
If this is the $j^\text{th}$ point in $\Pi_a$,
the new label is $aj$, the (potential) birth time is $\tau_{aj} = \tau_a + s$.
For convenience define
$$
    \ell(x, y, \eta, v)
    =
    \frac{
        v
    }{
        2 \lambda^{-1} \theta \gamma(x, \eta) r(x + y, \eta)
    } ,
$$
and using this, the (potential) new level is
\begin{equation*}
    U_{aj}(\tau_{aj})
    =
    U_a(\tau_{aj})
    +
    \ell(X_a(\tau_{aj}-), y, \eta(\tau_{aj}-), v) .
\end{equation*}
If $U_{aj}(\tau_{aj}) < \lambda$,
the new individual labeled $aj$ is produced,
and $\kappa$ determines which label is associated with the new location,
so
\begin{align*}
    X_{aj}(\tau_a + s)
    &=
    X_a(\tau_a+s-) + (1 - \kappa) y .
\end{align*}
On the other hand
if $U_{aj}(\tau_a+s) \ge \lambda$, then $X_a$ is unchanged and $X_{aj}$ is undefined,
so
\begin{align} \label{eqn:X_line_of_descent}
    X_a(\tau_a+s)
    &=
    X_a(\tau_a+s-) + \kappa y \ind_{U_{aj}(\tau_{aj}) < \lambda} .
\end{align}
Recall that the parental \emph{individual} always retains their spatial location,
so that $\kappa = 0$ corresponds to the parent being assigned a new level,
and our line of descent switching to the offspring.

@@

\comment{
    The following is quite annoying to get right, but can fixed up.
    However, I don't think we need it.}

Consequently, 
the \emph{line of descent}
leading from the initial generation
to any label $a \in \labelspace$
with $U_a(\tau_a) < \lambda$
can be explicitly constructed as follows.
Supposed that $a = (a_1, a_2, \ldots, a_n) \in \IN^n$,
and for $1 \le k \le n$ let $(a)_k = (a_1, a_2, \ldots, a_k)$.
Then, let $\zeta_a(t)$ denote the counting process that records
which label is present in the line of descent at time $t$,
i.e., $\zeta_a(t) = \max\{k \ge 1 : \tau_{(a)_k} \le t\}$.
This then lets us define $X_a(t)$ all the way back to $t=0$
by setting $X_a(t) = X_{(a)_{\zeta_a(t)}}(t)$ (and, $U_a(t)$ similarly).
Now, we can explicitly write down the evolution of $U_a(t)$ and $X_a(t)$.

First, for each $1 \le k < n$
let $(s_k, v_k, y_k, \kappa_k)$ be the $a_{k+1}^\text{th}$ point in $\Pi_{(a)_k}$
(this is the sequence of jumps taken by the line of descent).
Note that $s_k = \tau_{(a)_k}$, and also set $s_0 = 0$.
Let $\Gamma_a$ by the point process ``experienced by the line of descent leading to $a$'',
which is for $A \subset [0,\infty)^2 \times \IR^d \times \{0,1\}$
\comment{This is wrong, because $s$ is relative to birth time.}
\begin{align*}
\Gamma_a(A)
    &=
    \sum_{k=1}^n
    \Pi_{(a)_k}\left(A \cap [s_{k-1}, s_k) \times [0, \infty) \times \IR^d \times \{0,1\}\right) 
    \\ &\qquad {}
    + \sum_{k=1}^{n-1}
      \delta_{(s_k, v_k, y_k, 1 - \kappa_k)}(A) .
\end{align*}
Note that we have included the points at which the line of descent's level jumps,
flipping $\kappa$ at these points,
and as a result $\Gamma_a$ has the same distribution as $\Pi_a$
restricted to $[0, \sigma_a)$.
Then for $0 \le t \le \sigma_a$,
\begin{align*}
    X_a(t)
    &=
    X_{a_1}(0)
    +
    \int_{
        [0, t) \times [0, \infty) \times \IR^d \times \{0, 1\}
    }
    \kappa y
    \ind_{
        U_{(a)_{\zeta(s)}}(s) + \ell( X_{(a)_{\zeta(s)}}(s-), y, \eta(s-), v) < \lambda
    }
    \Gamma_{a}(ds, dv, dy, d\kappa) 
\end{align*}
and
\begin{align*}
    U_a(t)
    &=
    U_{a_1}(0)
    +
    \sum_{k=1}^{\zeta_a(t) - 1}
    \frac{
        v_k
    }{
        2 \lambda^{-1} \theta \gamma(X_{(a)_k}(s_k-), \eta(s_k-))
        r(X_{(a)_k}(s_k-) + y_k, \eta(s_k-))
    }
    \\ & \qquad {}
    +
    \int_0^t
    \bigg(
        \frac{\theta}{\lambda} \gamma(X_a(s),\eta(s))
        \int_{\IR^d} r(z,\eta(s)) q(X_a(s),dz) U_a(s)^2
    \\ & \qquad\qquad  \qquad {}
        -
        b(X_a(s),\eta(s)) U_a(s)
    \bigg)
    dt .
\end{align*}

@@

\comment{This is a rewrite of \ref{sec:lines_of_descent} following on the previous section}

@@

Now, let $\beta(x, \eta) = \meanq + 2 \Sigma \grad \log r(x, \eta)$,
and associate with each label an independent $d$-dimensional Brownian motion $W_a$,
and let $\covq = K K^T$.
Then, we can summarize the results above by saying that
in the infinite density limit,
the spatial locations and level of each line of descent $a$
behave for $\tau_a \le t < \sigma_a$ as
\begin{align*}
X_a(t)
    &=
    X_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \beta(x, \eta(s)) ds
        +
        K dW_a(s)
    \bigg)
    ds  \\
U_a(t)
    &=
    U_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \alpha \gamma(X_a(s),\eta(s))
        r(x,\eta(s)) U_a(s)^2
\\ &\qquad \qquad \qquad {}   
        -
        \left\{
            \gamma(X_a(s),\eta(s)) \DG r(X_a(s),\eta(s))
            + F(X_a(s), \eta(s))
        \right\}
        U_a(s)
    \bigg)
    ds .
\end{align*}
To complete the description,
for each label $a \in \labelspace$ let $R_a$ denote an independent Poisson process on $[0, \infty)^2$
with Lebesgue mean measure,
and as before order the points of each in some way.
If the population state at time 0 is $\eta(0)$,
then let $(x_i, u_i)$ be the points of a Poisson process
with mean measure on $\IR^d \times [0, \infty)$ a product of $\eta(0)$ and Lebesgue measure.
\comment{Here I've switched $t$ to be the absolute birth time
rather than relative to the birth time of the parent; I think it's clearer?
It, should be made to agree with the previous section.}
For each point begin a line of descent
with label $i$, location $X_i(0) = x_i$, level $U_a(0) = u_i$, and birth time $\tau_i = 0$.
Each point in the $R_a$ denotes a potential birth time for $a$:
if the $j^\text{th}$ point in $R_a$ is $(t, v)$,
$\tau_a \ge t$, and and $U_a(t) < \infty$,
then a new line of descent with label $aj$ is produced,
with birth time $\tau_{aj} = t$, location $X_{aj}(t) = X_a(t)$, and level
$$
    U_{aj}(t) = U_a(t)
    + \frac{v}{ 2 \alpha \gamma(X_a(t), \eta(t)) r(X_a(t), \eta(t)) } .
$$
\comment{Perhaps the conclusion of the Propostion would be something like:}
We can now express our population processes in the limit using these levels:
\begin{align*}
    \eta_t = \sum_{a : \tau_a > t \;\&\; U_a(t) < \lambda} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp = \sum_{a : \tau_a > t \;\&\; U_a(t) < \lambda} \delta_{(X_a(t), U_a(t))} .
\end{align*}

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{PREVIOUS VERSION OF Evolution of an Individual Particle}
\label{sec: Evolution of an Individual Particle}

Here we give a general framework of tracking evolution
of individual particles in the spatial-level population $\xi(t)_{t \geq 0}$ defined in \S \ref{sec:lookdown_defn}.

Particles alive at time zero will be indexed by $(0, i)$,
or more precisely,
we index the levels of the particles
alive at time $0$ by $(0, i)$,
keeping in mind that at each birth event,
the parent particle will,
with probability $1/2$, change levels.
Since no two particles born after time $0$
will be born at the same time,
we index the rest of the levels by their birth times.

>>>>

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{C}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels

@@

\comment{TODO: insert remark about writing $\gamma(x, \smooth{\gamma}\eta(x)) = \gamma(x, \eta)$.}

@@

\begin{proposition}
    \label{prop:nonlocal_to_local}
    Suppose that $\eta_\epsilon(t,dx)$ is a solution to equation \eqref{eqn:nonlocal_pde}
    with $r(x, m) = 1$ for all $x$ and $m$
    and some $\epsilon_r < \epsilon_\gamma = \epsilon_F = \epsilon$
    \comment{and assumptions on $\gamma$? and $F$?}.
    Then $\eta_\epsilon(t, dx)$ converges \comment{in what sense?} as $\epsilon \to 0$
    to $u(t, x) dx$, which solves
    \begin{align} \label{eqn:PDE}
        \partial_t u(t, x)
        =
        \DG^* \left( \gamma(\cdot, u(t, \cdot) u(t, \cdot)  \right)(x)
        + F(x, u(t, x)) u(t, x) .
    \end{align}
\end{proposition}

@@

\comment{Make forwards-backwards time notation reasonable in this.}

\comment{Perhaps write generator as depending on $\phi$ rather than $s$?
    Then we can say that the generator is $\Lgen_{\varphi(T-s)}$.}

\begin{theorem} \label{thm:lineages}
    In the context of Definition~\ref{def:lineage_generator},
    \begin{align}
        \label{eqn:lineage_generator}
        \Lgen_s f
        &=
        \frac{r}{\varphi_{T-s}}
        \left[
            \DG^*(\gamma \varphi_{T-s} f) 
            - f \DG^* (\gamma \varphi_{T-s})
        \right] \\
        &= \label{eqn:lineage_generator2}
        r\gamma
        \left[
            \sum_{ij} \covq_{ij} \partial_{ij} f
            + \sum_j \vec{m}(s)_j \partial_j f
        \right] ,
    \end{align}
    where $\vec{m}$ is the vector
    $$
    \vec{m}(s)_j
    =
    2 \sum_i C_{ij} \partial_i \log(\gamma \varphi_{T-s})
    + 2 \sum_i \partial_i C_{ij}
    - \meanq_j .
    $$
\end{theorem}

@@

Although our proof of Theorem~\ref{thm:lineages}
uses the explicit representation in terms of the lookdown process,
informal calculations agree with the main result.
Suppose that we are tracking a lineage that is at location $y$ at time $t$.
Looking back through time, the lineage will move at the time that individual was born,
and will jump to the location of the parent of that individual.
Now, the rate at which new individuals are born to parents at $x$ and establish at $y$
is
$$
    \gamma(x, \eta_t) q(x, dy) r(y, \eta_t) .
$$
Happily, we are interested in the case when $\eta$ has a density,
say $\eta_t(dx) = \varphi_t(x) dx$,
so interpreting these rates is not too difficult.
This implies that the probability a randomly chosen individual near $y$
is a new offspring from a parent at $x$ in $[t, t+dt)$ is, informally,
\begin{equation} \label{eqn:informal_rates}
\frac{
    \gamma(x, \eta_t) q_\theta(x, dy) r(y, \eta_t) \varphi_t(x)
}{
    \varphi_t(y)
} \frac{ q_\theta(x, dy) }{ dy } dx dt .
\end{equation}
Leaving aside questions of whether a lineage can be treated as a randomly chosen individual,
we might then proceed by defining a continuous-time jump process
whose transition rates are given by \eqref{eqn:informal_rates} given $(\varphi_t)_{t=0}^T$.
Let $(L^\theta_s)_{s=0}^T$ be the location of a lineage that moves according to
jump rates~\eqref{eqn:informal_rates},
back through time (i.e., with the substitution that $s = T-t$).
Then, writing $q_\theta(x, y)$ for the density of $q_\theta$,
\begin{align} \label{eqn:lineage_generator_Edt}
    \begin{split}
    &\IE[f(L^\theta_{s+ds}) - f(y) \;|\; L_s = y]
    \\&\qquad 
    =
    ds \int \left(f(x) - f(y)\right)
    \frac{
        \varphi_{T-s}(x) \gamma(x, \eta_{T-s}) r(y, \eta_{T-s})
    }{
        \varphi_{T-s}(y)
    }
    q_\theta(x, y) dx .
    \end{split}
\end{align}
\comment{I'm not sure if at this point in the paper we want to either
(a) write out the next statement; (b) restrict to the $\DG = \Delta$ case,
or (c) leave it as is (as an exercise for the reader, I guess?}
Referring back to Definition~\ref{def:dispersal_generator},
it is not too hard to show that as $\theta \to \infty$,
\begin{align*}
    \theta \int (f(x) - f(y)) q_\theta(x, y) dx
    &\to
    \DG^* f(x)  - f(x) \DG^* 1 .
\end{align*}
Consequently,
\begin{align*}
    &
    \theta \int (f(x) - f(y)) g(x) q_\theta(x, y) dx \\
    &\qquad =
    \theta \int \left\{
        (f(x) g(x) - f(y) g(y)) - f(y) (g(x) - g(y))
    \right\} q_\theta(x, y) dx \\
    &\qquad \to
        \DG^*(fg)(x) - f(x) \DG^* g(x) . 
\end{align*}
Applying this to~\eqref{eqn:lineage_generator_Edt} with $g = \varphi_{T-s} \gamma$,
this suggests that the generator of the limiting process is
\begin{align} \label{eqn:heuristic_lineage_generator}
    \Lgen_s f
    &=
    \frac{r}{\varphi_{T-s}}
    \left\{
        \DG^*(\gamma \varphi_{T-s} f) - f \DG^*(\gamma \varphi_{T-s})
    \right\} .
\end{align}
This agrees with Theorem~\ref{thm:lineages}.

@@

\comment{Maybe we can not introduce $\theta$ before this point in this section, and say we'll put it in now?}

@@

\comment{TODO: this note goes much earlier.}
To scale our lookdown model to be consistent with
the spatial model described in \S \ref{sec: Model and main results},
we will let $\lambda$ takes on the role of $N$
in \S \ref{sec: Model and main results},
i.e. each individual has mass $1/\lambda$
in the scaling limit.
Time is once again scaled by $\theta$,
and the birth, death and establishment rates 
depend on local densities given by
mollification over Gaussian kernels with radius
$\epsilon_{\gamma}$, $\epsilon_{F}$ and $\epsilon_{r}$.

@@

@@

Concretely, then: for each label $a$ in
$\labelspace = \bigcup_{k \ge 1} \IN_{\ge 1}^k$
let $\Pi_a$ be an independent Poisson process on $[0, \infty)^2 \times \IR_d \times \{0,1\}$.
The mean measure is a product of Lesbegue measure on the first two coordinates,
$q(0, \cdot)$ on $\IR^d$, and $(\delta_0 + \delta_1)/2$ on $\{0, 1\}$.
It will also be convenient
to suppose that for each label $a$ we have an enumeration of the points in $\Pi_a$,
so we may refer to ``the $j^\text{th}$ point in $\Pi_a$'',
although the precise order of this enumeration is irrelevant.
If $(t, v, y, \kappa)$ is the $j^\text{th}$ point in $\Pi_a$,
then $t$ will determine possible birth times,
$v$ will determine levels of the offspring,
$y$ will give the spatial displacement of the offspring relative to the parent,
$\kappa$ will be used to determine whether parent or offspring are assigned the new level,
and the new label produced will be $a \concat j$,
i.e., the label $a$ with $j$ appended (so, if $a \in \IN_{\ge 0}^k$ then $aj \in \IN_{\ge 0}^{k+1}$).
Each label $a$ has a birth time $\tau_a$,
when it is first assigned,
and a death time $\sigma_a$, when its level first hits $\lambda$.
For any $\tau_a \le t \le \sigma_a$ we denote by $X_a(t)$ and $U_a(t)$ the spatial location and level
of the individual carrying label $a$ at time $t$.
Now, since we have defined labels so that the level does not jump,
$U_a$ satistifies~\eqref{eqn:dot_u} for $\tau_a \le t \le \sigma_a$, i.e.,
\begin{equation} \label{eqn:U_line_of_descent}
    \begin{split}
& U_a(t)
    =
    U_a(\tau_a) \\
&\qquad {}   
    + \int_{\tau_a}^{t}
    \left(
        \frac{\theta}{\lambda} \gamma(X_a(s),\eta(s))
        \int_{\IR^d} r(z,\eta(s)) q(X_a(s),dz) U_a(s)^2
        -
        b(X_a(s),\eta(s)) U_a(s)
    \right)
    ds ,
\end{split}
\end{equation}
and, of course, $\sigma_a = \inf\{t \ge \tau_a : U_a(t) > \lambda\}$.

Potential reproduction events occur at times $\tau_a + s$
for each point $(s, v, y, \kappa) \in \Pi_a$.
We say `potential' since if the resulting level is greater than $\lambda$,
the event does not happen.
If this is the $j^\text{th}$ point in $\Pi_a$,
the new label is $a \concat j$, the (potential) birth time is $\tau_{a \concat j} = \tau_a + s$.
For convenience define
$$
    \ell(x, y, \eta, v)
    =
    \frac{
        v
    }{
        2 \lambda^{-1} \theta \gamma(x, \eta) r(x + y, \eta)
    } ,
$$
and using this, the (potential) new level is
\begin{equation*}
    U_{a \concat j}(\tau_{a \concat j})
    =
    U_a(\tau_{a \concat j})
    +
    \ell(X_a(\tau_{a \concat j}-), y, \eta(\tau_{a \concat j}-), v) .
\end{equation*}
If $U_{a \concat j}(\tau_{a \concat j}) < \lambda$,
the new individual labeled $a \concat j$ is produced,
and $\kappa$ determines which label is associated with the new location,
so
\begin{align*}
    X_{a \concat j}(\tau_a + s)
    &=
    X_a(\tau_a+s-) + (1 - \kappa) y .
\end{align*}
On the other hand
if $U_{a \concat j}(\tau_a+s) \ge \lambda$, then $X_a$ is unchanged and $X_{a \concat j}$ is undefined,
so
\begin{align} \label{eqn:X_line_of_descent}
    X_a(\tau_a+s)
    &=
    X_a(\tau_a+s-) + \kappa y \ind_{U_{a \concat j}(\tau_{a \concat j}) < \lambda} .
\end{align}
Recall that the parental \emph{individual} always retains their spatial location,
so that $\kappa = 0$ corresponds to the parent being assigned a new level,
and our line of descent switching to the offspring.

@@

\begin{remark}
    \comment{TODO: say that it's clear how this works following a line of descent
    all the way from time 0 (in particular, in the infinite density limit,
    just the spatial part looks just the same despite the jumps in level).}
\end{remark}

@@

@@

Now, let $\beta(x, \eta) = \meanq + 2 \Sigma \grad \log r(x, \eta)$,
and associate with each label an independent $d$-dimensional Brownian motion $W_a$,
and let $\covq = K K^T$.
Then, we can summarize the results above by saying that
in the infinite density limit,
the spatial locations and level of each line of descent $a$
behave for $\tau_a \le t < \sigma_a$ as
\begin{align*}
X_a(t)
    &=
    X_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \beta(x, \eta(s)) ds
        +
        K dW_a(s)
    \bigg)
    ds  \\
U_a(t)
    &=
    U_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \alpha \gamma(X_a(s),\eta(s))
        r(x,\eta(s)) U_a(s)^2
\\ &\qquad \qquad \qquad {}   
        -
        \left\{
            \gamma(X_a(s),\eta(s)) \DG r(X_a(s),\eta(s))
            + F(X_a(s), \eta(s))
        \right\}
        U_a(s)
    \bigg)
    ds .
\end{align*}
To complete the description,
for each label $a \in \labelspace$ let $R_a$ denote an independent Poisson process on $[0, \infty)^2$
with Lebesgue mean measure,
and as before order the points of each in some way.
If the population state at time 0 is $\eta(0)$,
then let $(x_i, u_i)$ be the points of a Poisson process
with mean measure on $\IR^d \times [0, \infty)$ a product of $\eta(0)$ and Lebesgue measure.
\comment{Here I've switched $t$ to be the absolute birth time
rather than relative to the birth time of the parent; I think it's clearer?
It, should be made to agree with the previous section.}
For each point begin a line of descent
with label $i$, location $X_i(0) = x_i$, level $U_a(0) = u_i$, and birth time $\tau_i = 0$.
Each point in the $R_a$ denotes a potential birth time for $a$:
if the $j^\text{th}$ point in $R_a$ is $(t, v)$,
$\tau_a \ge t$, and and $U_a(t) < \infty$,
then a new line of descent with label $a \concat j$ is produced,
with birth time $\tau_{a \concat j} = t$, location $X_{a \concat j}(t) = X_a(t)$, and level
$$
    U_{a \concat j}(t) = U_a(t)
    + \frac{v}{ 2 \alpha \gamma(X_a(t), \eta(t)) r(X_a(t), \eta(t)) } .
$$
\comment{Perhaps the conclusion of the Propostion would be something like:}
We can now express our population processes in the limit using these levels:
\begin{align*}
    \eta_t = \sum_{a : \tau_a > t \;\&\; U_a(t) < \lambda} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp = \sum_{a : \tau_a > t \;\&\; U_a(t) < \lambda} \delta_{(X_a(t), U_a(t))} .
\end{align*}

\comment{TODO: incorporate this in the discussion somewhere (and maybe other refs? see previous version).}
This representation of individuals allows us to identify all 
of the individuals in the current population that are descendants of a
given ancestor at time zero. In theory at least, this allows us to
recover all the information about genealogies relating individuals 
sampled from the present day population. This idea draws on the notion
of `tracers', popular in statistical physics and used in population
genetics by a number of authors including 
\cite{biswas/etheridge/klimek:2018, durrett/fan:2016, hallatschek/nelson:2008}.
